---
title: "PPOL 670, Assignment 04"
author: "Erin Lemons"
date: "10/3/2021"
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

\begin{center}
\Huge{PPOL 670 | Assignment 4}

\Huge{Data Visualization, Markdown, and Git}
\end{center}

\vspace{0.1in}

```{r library}
library(tidyverse)
library(ggplot2)
library(haven)
options(scipen = 999)
library(ggthemes)
```

# Data
Data throughout this assignment is from the V-Dem Dataset - Version 11.1 located at https://www.v-dem.net/en/data/data/v-dem-dataset-v111/#.

Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, Nazifa
Alizada, David Altman, Michael Bernhard, Agnes Cornell, M. Steven Fish, Lisa Gastaldi,
Haakon Gjerløw, Adam Glynn, Allen Hicken, Garry Hindle, Nina Ilchenko, Joshua Krusell,
Anna L¨uhrmann, Seraphine F. Maerz, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova,
Juraj Medzihorsky, Pamela Paxton, Daniel Pemstein, Josefine Pernes, Johannes von R¨omer,
Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundstr¨om, Eitan
Tzelgov, Yi-ting Wang, Tore Wig, Steven Wilson and Daniel Ziblatt. 2021. ”V-Dem [Country–
Year/Country–Date] Dataset v11.1” Varieties of Democracy (V-Dem) Project. https://doi.
org/10.23696/vdemds21.

Pemstein, Daniel, Kyle L. Marquardt, Eitan Tzelgov, Yi-ting Wang, Juraj Medzihorsky, Joshua
Krusell, Farhad Miri, and Johannes von R¨omer. 2021. “The V-Dem Measurement Model: Latent
Variable Analysis for Cross-National and Cross-Temporal Expert-Coded Data”. V-Dem
Working Paper No. 21. 6th edition. University of Gothenburg: Varieties of Democracy Institute.

# The Issue: Government Internet Shutdowns
Governments all over the world have resorted to Internet shutdowns in (often unsuccessful) attempts to quell domestic unrest and even to prevent cheating on national exams since President Lansana Conte first tried this approach in 2007 by ordering all of Guinea's Internet service providers to suspend services in response to domestic protests (Rydzak et al. 2020). Internet shutdowns have negative economic and social impacts and often disproportionately affect the world's most vulnerable populations (https://www.accessnow.org/keepiton/#problem). Comparative government scholars are still seeking to better understand this phenomenon in order to develop policy solutions to combat this problem.  Specifically, under what conditions do states resort to Internet shutdowns?

# The Threat: Government Shutdown Capacity


# Is this about Regime Type?

# Is this about State Capacity?

# An Indicator of Relative Weakness?
