---
title: "PPOL 670, Assignment 04"
author: "Erin Lemons"
date: "10/3/2021"
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

\begin{center}
\Huge{PPOL 670 | Assignment 4}

\Huge{Data Visualization, Markdown, and Git}
\end{center}

\vspace{0.1in}

```{r library}
library(tidyverse)
library(ggplot2)
library(haven)
options(scipen = 999)
library(ggthemes)
```

# Data
Data throughout this assignment is from the V-Dem Dataset - Version 11.1 located at https://www.v-dem.net/en/data/data/v-dem-dataset-v111/#.

Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, Nazifa
Alizada, David Altman, Michael Bernhard, Agnes Cornell, M. Steven Fish, Lisa Gastaldi,
Haakon Gjerløw, Adam Glynn, Allen Hicken, Garry Hindle, Nina Ilchenko, Joshua Krusell,
Anna L¨uhrmann, Seraphine F. Maerz, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova,
Juraj Medzihorsky, Pamela Paxton, Daniel Pemstein, Josefine Pernes, Johannes von R¨omer,
Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundstr¨om, Eitan
Tzelgov, Yi-ting Wang, Tore Wig, Steven Wilson and Daniel Ziblatt. 2021. ”V-Dem [Country–
Year/Country–Date] Dataset v11.1” Varieties of Democracy (V-Dem) Project. https://doi.
org/10.23696/vdemds21.

Pemstein, Daniel, Kyle L. Marquardt, Eitan Tzelgov, Yi-ting Wang, Juraj Medzihorsky, Joshua
Krusell, Farhad Miri, and Johannes von R¨omer. 2021. “The V-Dem Measurement Model: Latent
Variable Analysis for Cross-National and Cross-Temporal Expert-Coded Data”. V-Dem
Working Paper No. 21. 6th edition. University of Gothenburg: Varieties of Democracy Institute.

```{r import data}
VDEM <- read_csv("V-Dem-CY-Full+Others-v11.1.csv")
```

# The Issue: Government Internet Shutdowns
Governments all over the world have resorted to Internet shutdowns in (often unsuccessful) attempts to quell domestic unrest and even to prevent cheating on national exams since President Lansana Conte first tried this approach in 2007 by ordering all of Guinea's Internet service providers to suspend services in response to domestic protests (Rydzak et al. 2020). Internet shutdowns have negative economic and social impacts and often disproportionately affect the world's most vulnerable populations (https://www.accessnow.org/keepiton/#problem). Comparative government scholars are still seeking to better understand this phenomenon in order to develop policy solutions to combat this problem.  Specifically, under what conditions do states resort to Internet shutdowns?

# The Threat: Government Shutdown Capacity
```{r first graph}
VDEM %>%
  filter(year >= 2015) %>%
  ggplot() +
  geom_histogram(mapping = 
                   aes(x= v2smgovshutcap_mean), 
                 binwidth = 0.1, 
color = "black", fill= "white") +
  scale_x_continuous(labels=c("No Capacity", "1/4 of Internet", "1/2 of Internet", "3/4 of Internet", "All Interent")) +
  geom_vline(mapping = 
               aes(
                 xintercept= 
                   mean(
                   x= v2smgovshutcap_mean)),
            color="blue", 
            linetype="dashed", 
            size=1) +
  annotate("text", x=2.5 ,y=55,label="mean",hjust=0, size=3, angle=90, color="blue") +
  labs(title="Most Governments Can Shut Down Their Domestic Internet", subtitle = "Year: 2015", x="Shutdown Capacity", y = "Count", caption = "Source: V-Dem Dataset") +
  theme_clean()+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), axis.text.x= element_text(size = 6))
```


In 2015, the vast majority of governments around the world had the capacity to shut down at least 1/4 of their domestic Internet, and on average, governments could shut down between 1/2 and 3/4 of their domestic Internet coverage. Despite their technical capacity, the overwhelming majority of governments do not shut down their Internet. From July 2015-June 2016, only 19 countries shut down parts of their domestic Internet for a combined total of 81 times (West, Darrell, "Internet Shutdowns Cost Countries $2.4 Billion Last Year," Center for Technology Innovation at Brookings, October 2016). What type of governments make this costly decision?

# Is this about Regime Type?
```{r second graph}
VDEM %>%
  filter(year == 2015) %>%
  mutate(inverse_v2smgovshut_mean= v2smgovshut_mean*-1) %>%
  filter(!is.na(x= v2x_regime)) %>%
  ggplot() +
  geom_boxplot(mapping = aes (x = factor(v2x_regime), y= inverse_v2smgovshut_mean, fill = factor(v2x_regime))) +
  scale_fill_manual(values = c("light grey", "light grey", "blue", "light grey")) +
  scale_x_discrete(labels=c("Closed Autocracy", "Electoral Autocracy", "Electoral Democracy", "Liberal Democracy")) +
  scale_y_continuous(labels = c("Never", "Rarely", "Somtimes", "Often")) +
  labs(title="More Democratic States Shut Down Internet Less Often", subtitle = "Year: 2015", x="Regime Type", y = "Internet Shutdowns", caption = "Source: V-Dem Dataset") +
  theme_clean()+
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), axis.text.x= element_text(size = 7), axis.text.y= element_text(size =7, angle =50))
```

The free Internet is has become closely associated wtih the concepts of freedom of expression and democracy (e.g., Warf 2011, Bak et al. 2018, Krcmaric 2019). Indeed, as the above graph demonstrates, there is certainly a correlation between regime type and Internet shutdown propensity. Nevertheless, there is a decent amount of variation within the same regime type.  Especially, so-called "electoral democracies" make it clear that just because a state is a "democracy," does not mean that it will never resort to an Internet shutdown.  In fact, India, a democracy, shuts down the Internet more frequently than any other country in the world ("#KeepItOn Report, Accessnow.org, 3 March 2021).


# Is this about State Capacity?
```{r third graph}

```

# An Indicator of Relative Weakness?
```{r fourth graph}

```
